Microbiome and hypertension

### 2018-12-20 rrbiome.Rmd
First maaslin calculation

library(dplyr)
library(tibble)
library(phyloseq)
library(microbiome)
library(finriskmetagcommon)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(ggpmisc)

@@

```{r adonis}

if (!file.exists("rds/phf.adonis.rds")) {
    phf.adonis <- adonis(dist.matrix ~ BL_AGE + SEX + MAP,
                         data = meta(pseq_filtered),
                         permutations = 99)
    saveRDS(phf.adonis, file = "rds/phf.adonis.rds")
} else {
    phf.adonis <- readRDS("rds/phf.adonis.rds")
}

phf.adonis$aov.tab %>% as.data.frame %>% kable

# PCA

```{r pca}
pseq.clr <- microbiome::transform(pseq_filtered, "clr")
X <- as.matrix(t(abundances(pseq.clr)))
prcomp <- stats::prcomp(X)
```

@@

>>>>

library(dplyr)
library(tibble)
library(phyloseq)
library(microbiome)
library(finriskmetagcommon)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(ggpmisc)
library(Maaslin)

@@

```{r adonis1}
adonis(dist.matrix ~ MAP, 
       data = meta(pseq_filtered),
       permutations = 99)$aov.tab %>%
                            as.data.frame %>%
                            kable
```

```{r adonis2}
adonis(dist.matrix ~ BL_AGE + SEX + MAP,
       data = meta(pseq_filtered),
       permutations = 99)$aov.tab %>%
                            as.data.frame %>%
                            kable
```

```{r adonis3}
adonis(dist.matrix ~ MAP + BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + BP_TREAT,
       data = meta(pseq_filtered),
       permutations = 99)$aov.tab %>%
                            as.data.frame %>%
                            kable
```

@@

# Genus

Aaron work-december-2018 prepare_maaslin_input.R

```{r prepare maaslin}

pheno <- t(meta(sample_data(pseq_filtered)))
sample.ids <- colnames(pheno)
phfg.pseq.rel <- microbiome::transform(pseq_filtered, "compositional")
abud.rel <- abundances(phfg.pseq.rel)

rownames(abud.rel) <- sapply(rownames(abud.rel),
                             function (genus) gsub("(.*) \\(+(.+)\\)", '\\1.\\2', genus))

merged <- rbind(sample.ids, pheno, abud.rel)
write.table(t(merged), file="merged_phfinrisk_genus.tsv", sep='\t', 
            quote=FALSE, row.names=FALSE, col.names=TRUE)

variables <- rownames(pheno)
variables_wo_ctrls <- variables[!variables %in% c("BL_AGE","BMI","SEX")]
data <- file("maaslin_config.config" , open='wt')
pcl.contents <- c("Matrix: Metadata",
                  paste(c("Read_PCL_Rows: BL_AGE,BMI,SEX",variables_wo_ctrls), collapse=","),
                  "",
                  "Matrix: Adundance",
                  "Read_PCL_Rows: Archaea.Archaea-")
writeLines(pcl.contents, con=data)
close(data)
```

```{r run maaslin, eval = FALSE} 
Maaslin("merged_phfinrisk_genus.tsv",
        "maaslin",
        strInputConfig = "maaslin_config.config",
        strModelSelection="none",
        dSignificanceLevel = 1.0, 
        fAllvAll = TRUE,
        strForcedPredictors = c("BL_AGE", "BMI", "SEX"))

Data framen luku *comp_maaslin_results_from_single_job.R* ja plottaus käyttäen *comp_maaslin_plots.R* ja *edgeR_plot_functions.R*.

```{r Reading maaslins output}

maas.results.df <- read.table("maaslin/merged_phfinrisk_genus.txt", header=TRUE, sep='\t')
maas.results.df[,'Covariate'] <- maas.results.df[,'Variable']
maas.results.df <- merge(maas.results.df,
                         names.dset[,c('Covariate','Category', 'Name')],
                         by=c('Covariate'))

maas.results.df[,'Maaslin.Q.value'] <- maas.results.df[,'Q.value']
maas.results.df[,'Qval'] <- p.adjust(maas.results.df[,'P.value'], method="BH")
maas.results.df[,'beta'] <- maas.results.df[,'Coefficient']
maas.results.df.signf <- dplyr::filter(maas.results.df, Qval < 0.05)

```{r scatter plots}
sysm.signf.features <- maas.results.df.signf %>% dplyr::filter(Covariate == "SYSTM") %>% pull(Feature)
```

```{r maaslin plots, fig.width = 10, fig.height = 4, dpi = 300}
source("edgeR_plot_functions.R")

plot_maaslin_heatmap_final(maas.results.df.signf,
                      xlab.text="Genera",
                      beta.maxabs=max(abs(maas.results.df.signf[,'beta'])))
